Forecasting the 2024 U.S. Presidential Election

### 2024-11-03 paper/paper.qmd
changed premable on all documents as well as adding llm usage, and updated readme file

# # Melt data for easier plotting
# prediction_melted <- melt(
#   prediction_comparison, 
#   id.vars = c("state", "electoral_votes", "winner"),
#   measure.vars = c("predicted_pct_harris", "predicted_pct_trump")
# )
# 
# # Rename melted variables for readability
# colnames(prediction_melted)[colnames(prediction_melted) == "variable"] <- "candidate"
# prediction_melted$candidate <- ifelse(
#   prediction_melted$candidate == "predicted_pct_harris", 
#   "Harris", 
#   "Trump"
# )
# 
# # Calculate the midpoint to split the states
# n_states <- length(unique(prediction_melted$state))
# midpoint <- ceiling(n_states / 2)
# 
# # Order states by electoral votes in descending order for better visualization
# states_order <- prediction_melted %>%
#   distinct(state, electoral_votes) %>%
#   arrange(desc(electoral_votes)) %>%
#   pull(state)
# 
# # Split the states into two groups
# left_states <- states_order[1:midpoint]
# right_states <- states_order[(midpoint + 1):length(states_order)]
# 
# # Create two separate datasets
# prediction_left <- prediction_melted %>%
#   filter(state %in% left_states) %>%
#   mutate(state = factor(state, levels = left_states))  # Ensure consistent ordering
# 
# prediction_right <- prediction_melted %>%
#   filter(state %in% right_states) %>%
#   mutate(state = factor(state, levels = right_states))  # Ensure consistent ordering
# 
# # Function to create individual plots to avoid repetition
# create_plot <- function(data, show_legend = FALSE) {
#   ggplot(data, aes(x = reorder(state, electoral_votes), y = value, fill = candidate)) +
#     geom_bar(
#       stat = "identity", 
#       position = position_dodge(width = 0.9), 
#       width = 0.5  # Reduced bar width for more spacing
#     ) +
#     coord_flip() +
#     labs(
#       x = NULL,
#       y = "Predicted Vote Percentage (%)"
#     ) +
#     scale_fill_manual(values = c("Harris" = "#87CEFA", "Trump" = "#F08080")) +
#     theme_minimal() +
#     theme(
#       text = element_text(size = 9),  # Reduced overall text size
#       axis.text.y = element_text(size = 8, margin = margin(r = 5)),  # Smaller state labels
#       axis.text.x = element_text(size = 9),
#       legend.title = element_blank(),
#       legend.position = ifelse(show_legend, "top", "none"),  # Place legend only for the first plot
#       legend.text = element_text(size = 9),
#       panel.grid.major.y = element_blank(),
#       plot.margin = margin(l = 10, r = 15, t = 10, b = 10)  # Adjusted margins
#     ) +
#     geom_text(
#       aes(label = round(value, 1)),
#       position = position_dodge(width = 0.9),
#       hjust = -0.2,    # Adjusted horizontal position for labels
#       size = 1.5,      # Smaller label size
#       show.legend = FALSE
#     ) +
#     scale_y_continuous(
#       expand = expansion(mult = c(0, 0.2)),  # Increased expansion for labels
#       limits = c(0, 75)  # Adjust based on your data
#     )
# }
# 
# # Create the left plot with legend
# plot_left <- create_plot(prediction_left, show_legend = TRUE)
# 
# # Create the right plot without legend
# plot_right <- create_plot(prediction_right, show_legend = FALSE)
# 
# # Combine the plots side by side with legend on top
# combined_plot <- plot_left + plot_right +
#   plot_layout(guides = "collect") &  # Collects the legend at the top
#   theme(legend.position = "top")  # Places the combined legend at the top
# 
# # Print the combined plot
# combined_plot

>>>>

# Melt data for easier plotting
prediction_melted <- melt(
  prediction_comparison, 
  id.vars = c("state", "electoral_votes", "winner"),
  measure.vars = c("predicted_pct_harris", "predicted_pct_trump")
)

# Rename melted variables for readability
colnames(prediction_melted)[colnames(prediction_melted) == "variable"] <- "candidate"
prediction_melted$candidate <- ifelse(
  prediction_melted$candidate == "predicted_pct_harris", 
  "Harris", 
  "Trump"
)

# Calculate the midpoint to split the states
n_states <- length(unique(prediction_melted$state))
midpoint <- ceiling(n_states / 2)

# Order states by electoral votes in descending order for better visualization
states_order <- prediction_melted %>%
  distinct(state, electoral_votes) %>%
  arrange(desc(electoral_votes)) %>%
  pull(state)

# Split the states into two groups
left_states <- states_order[1:midpoint]
right_states <- states_order[(midpoint + 1):length(states_order)]

# Create two separate datasets
prediction_left <- prediction_melted %>%
  filter(state %in% left_states) %>%
  mutate(state = factor(state, levels = left_states))  # Ensure consistent ordering

prediction_right <- prediction_melted %>%
  filter(state %in% right_states) %>%
  mutate(state = factor(state, levels = right_states))  # Ensure consistent ordering

# Function to create individual plots to avoid repetition
create_plot <- function(data, show_legend = FALSE) {
  ggplot(data, aes(x = reorder(state, electoral_votes), y = value, fill = candidate)) +
    geom_bar(
      stat = "identity", 
      position = position_dodge(width = 0.9), 
      width = 0.5  # Reduced bar width for more spacing
    ) +
    coord_flip() +
    labs(
      x = NULL,
      y = "Predicted Vote Percentage (%)"
    ) +
    scale_fill_manual(values = c("Harris" = "#87CEFA", "Trump" = "#F08080")) +
    theme_minimal() +
    theme(
      text = element_text(size = 9),  # Reduced overall text size
      axis.text.y = element_text(size = 8, margin = margin(r = 5)),  # Smaller state labels
      axis.text.x = element_text(size = 9),
      legend.title = element_blank(),
      legend.position = ifelse(show_legend, "top", "none"),  # Place legend only for the first plot
      legend.text = element_text(size = 9),
      panel.grid.major.y = element_blank(),
      plot.margin = margin(l = 10, r = 15, t = 10, b = 10)  # Adjusted margins
    ) +
    geom_text(
      aes(label = round(value, 1)),
      position = position_dodge(width = 0.9),
      hjust = -0.2,    # Adjusted horizontal position for labels
      size = 1.5,      # Smaller label size
      show.legend = FALSE
    ) +
    scale_y_continuous(
      expand = expansion(mult = c(0, 0.2)),  # Increased expansion for labels
      limits = c(0, 75)  # Adjust based on your data
    )
}

# Create the left plot with legend
plot_left <- create_plot(prediction_left, show_legend = TRUE)

# Create the right plot without legend
plot_right <- create_plot(prediction_right, show_legend = FALSE)

# Combine the plots side by side with legend on top
combined_plot <- plot_left + plot_right +
  plot_layout(guides = "collect") &  # Collects the legend at the top
  theme(legend.position = "top")  # Places the combined legend at the top

# Print the combined plot
combined_plot
